---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(tidyverse)
```

# ruijter

<!-- badges: start -->
<!-- badges: end -->

`{ruijter}` is an R data package that provides the real-time qPCR technical data
sets used in [Ruijter et al.
(2013)](https://doi.org/10.1016/j.ymeth.2012.08.011), namely:

- The 94-replicates-4-dilutions data set: `ds_94_4`
- The 380 replicates data set: `ds_380`
- The competimer data set: `ds_competimer`


## Installation

You can install the development version of `{ruijter}` like so:

``` r
# install.packages("remotes")
remotes::install_github("ramiromagno/ruijter")
```

## Usage


### Four-point 10-fold dilution series `ds_94_4`

```{r dpi=300, fig.width=8, fig.height=4}
library(ruijter)

head(ds_94_4)

dplyr::count(ds_94_4, well, replicate, sample_type, copies)

ds_94_4 %>%
  ggplot(mapping = aes(x = cycle, y = fluor, group = well, col = as.character(copies))) +
  geom_line(size = 0.1) +
  labs(y = "Raw fluorescence", colour="Copy number", title = "Four-point 10-fold dilution series") +
  guides(color = guide_legend(override.aes = list(size = 0.5)))
```

###  Replicates for assessment of precision `ds_380`

```{r dpi=300, fig.width=8, fig.height=4}
head(ds_380)

dplyr::count(ds_380, well, replicate, sample_type, copies)

ds_380 %>%
  ggplot(mapping = aes(x = cycle, y = fluor, group = well, col = as.factor(copies))) +
  geom_line(size = 0.1) +
  labs(y = "Raw fluorescence", colour="Copy number", title = "380 replicates data set") +
  guides(color = guide_legend(override.aes = list(size = 0.5)))
```

###  Competimer primers for PCR efficiency modulation `ds_competimer`

```{r dpi=300, fig.width=8, fig.height=8}
head(ds_competimer)

dplyr::count(ds_competimer, well, pct, conc, replicate, sample_type)

ds_competimer %>%
  ggplot(mapping = aes(x = cycle, y = fluor, group = interaction(pct, conc, replicate), col = interaction(pct, conc))) +
  geom_line(size = 0.2) +
  guides(color = "none") +
  labs(y = "Raw fluorescence", title = "Competimer data set") +
  facet_grid(rows = vars(pct), cols = vars(conc))
```

## Admonition

- The original data on which data set `ds_94_4` is based seems to have incorrect annotated the standards' copy numbers. The order seems to be reversed after visual inspection of the curves, e.g., the late curves, emerging around cycle 32, are annotated with 15,000 copies of starting material. This is totally the reverse expectation. So in `{ruijter}` we have fixed this issue by reversing the annotation:
  - `15` becomes `15,000`
  - `150` becomes `1,500`
  - `1,500` becomes `150`
  - `15,000` becomes `15`

- The data set `ds_380` which comprises (almost) of replicates at copy number
`15,000` also does not seem to meet the expectation given the curves of the same
dilution level in `ds_94_4`. The curves in `ds_380` rise earlier, and hence seem
to correspond to a higher concentration than those of `ds_94_4`. At the moment,
no correction has been performed but we will investigate this further to
understand if this discrepancy also corresponds to an obvious mistake.

## Terms of use

If you use the data here provided please do not forget to cite the original
work by Ruijter et al. (2013), and this package.

## Code of Conduct
  
Please note that the `{ruijter}` project is released with a [Contributor Code of
Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By
contributing to this project, you agree to abide by its terms.

## References

Jan M. Ruijter, Michael W. Pfaffl, Sheng Zhao, Andrej N. Spiess, Gregory Boggy,
Jochen Blom,Robert G. Rutledge, Davide Sisti, Antoon Lievens, Katleen De Preter,
Stefaan Derveaux, Jan Hellemans, Jo Vandesompele. *Evaluation of qPCR curve
analysis methods for reliable biomarker discovery: Bias, resolution, precision,
and implications*. Methods 59 32--46 (2013). doi:
[10.1016/j.ymeth.2012.08.011](https://doi.org/10.1016/j.ymeth.2012.08.011).
